---
title: "WUP_04"
author: "Michael Maida"
date: "fall 2021"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

We need to take a look at the data to see which helper file we will need. We will load the data and see what type of data we are given. 

```{r}
OurData<-read.csv(file="GSScsv.csv")
```
We see wordsum is a numeric variable and degree is a factor variable. This shows that we need to upload our helper file "numerical~categorical."


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(OurData,aes(x=DEGREE,y=WORDSUM))

basicCN + geom_boxplot()
```

The graduate's box is higher in the y-axis than the high school and lt high school box. It looks like there is a difference in the wordsum test.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=OurData)
```

The mean for lt high school is 7.75 while the mean for graduate school is 10.09.This indicates a meaningful difference between the two.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=OurData)
summary.aov(model1)
```

The P-value is 0.000236. This P-value represents the chance that we would get he data under the conditions that there isn't a difference between each degree. Since the P-value is so little, it looks as it there is indeed a difference.

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 16, 2, 2))
plot(TukeyHSD(model1),las=2)
```

There are two conditions where the score on the degree are statistically different, and they are between high school and graduate school and lt high school and graduate school. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

We looked at the problem and figured out it was a nemerical~categorical problem, so we uploaded the nemerical~categorical helper file. We looked at the different T-tests and figured out which confidence levels were less than the rest. We concluded that the conditions between high school and graduate school and between lt high school and graduate schoool were significantly different. 